---
title: 'RMarkdown: Titanic Case Study'
output: "html_document"
author: "Jian Wang"
date: "`r format(Sys.time(), '%B %d, %Y')`"
---


This is my first R Markdown document for **MSDS 593**. I am required to submit all **MSDS 593** homework in RMarkdown.

Firstly, I am going to generate an [html]() document as output, set in the YAML header section of this document.
Secondly, I am **NOT** hardcoding the date, but rather using a function which will automatically print the current date on the day the document is knit.

Now, I am going to import a dataset about passengers from the Titanic, using the following line of code:
```{r readData,eval=FALSE}
read.csv("/Users/jw/classes/msan593/Exploratory-Data-Analysis-with-R/courseData/titanic.csv")
```

This fails for a few reasons, namely, I read in the file and stored it no where. So I wasted my time waiting for R to read in the file, and then when it finally did, it printed the rows of data to the console window, and voila, the data disappered faster than it loaded. Now I knwo better.

```{r}
titanicData <- read.csv("/Users/jw/classes/msan593/Exploratory-Data-Analysis-with-R/courseData/titanic.csv")
```

titanicData has `r nrow(titanicData)` rows and `r ncol(titanicData)` columns. All variables in the data frame are listed here using str().
```{r printout, echo=FALSE}
str(titanicData)
```

The Titanic had `r length(titanicData$Sex[titanicData$Sex == "male"])` male passengers(`r round(length(titanicData$Sex[titanicData$Sex=="male"]) /nrow(titanicData)*100, digits=2)`%) and `r length(titanicData$Sex[titanicData$Sex =="female"])` female passengers (`r round(length(titanicData$Sex[titanicData$Sex=="female"]) /nrow(titanicData)*100, digits=2)`%).

The mean age of all passengers is `r format(mean(titanicData$Age,na.rm=T),digits=4)`.I was careful not to print output `r mean(titanicData$Age,na.rm=T)`, which has far too many numbers after the decimal. Also note that when using the `mean` function,one should not forget to include the `na.rm= TRUE` option, to ensure that an  `NA` result is not returned. The mean age of all male passengers is `r format(mean(titanicData$Age[titanicData$Sex=="male"],na.rm=T),digits=4)`, and the mean age of all female passengers is `r format(mean(titanicData$Age[titanicData$Sex=="female"],na.rm=T),digits=4)`.

I would like to figure out the breakdown of children versus adults on the Titanic. Firstly, I will create a new column in the `titanicData` data frame called `adultOrChild`, in which I will store either `child` or `adult` after testing the age of each passenger. 

```{r echo = TRUE}
  titanicData <- read.csv("/Users/jw/classes/msan593/Exploratory-Data-Analysis-with-R/courseData/titanic.csv")
  titanicData$adultOrChild <- as.factor(ifelse(titanicData$Age>=18,"adult","child"))
```


It turns out there were `r length(titanicData$adultOrChild[titanicData$adultOrChild=="child"])` children on the ship and `r length(titanicData$adultOrChild[titanicData$adultOrChild=="adult"])` adults. No age data people are `r sum(is.na(titanicData$Age))`.

People with their cabin information on record are `r length(titanicData$Cabin[titanicData$Cabin!= ""])` out of all `r nrow(titanicData)`(`r format(round(length(titanicData$Cabin[titanicData$Cabin!= ""])/nrow(titanicData)*100,digits=1))`%). 

Generate summary statistics on passenger fares:

```{r echo=F}
summary(titanicData$Fare)
```
Generate the histogram.
```{r echo=F}
hist(titanicData$Age,freq = T)
```


















