---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
Test all about type, class, and  mode.
```{r}
10352/987653 #
2^6*99
914545%%33
cat("\014")
```

kp1 : function masking, when different package contains same name for certain function, should you be careful. A good example is in library(Hmisc), the function summarize share the same name as that in library(dplyr), so when doing piping, use dplyr::summarize to make sure it works. 

About packages.
From installed.packages(), you will get installed.packages scans the ‘DESCRIPTION’ files of each package found along lib.loc and returns a matrix of package names, library paths and version numbers.
```{r eval=F}
#install package:
install.packages("dplyr")
#check installed package
installed.packages()
#see all loaded package
search()
(.packages())
#load the package
library(tidyverse)
#detach a package
detach("package:tidyverse",unload=T)

```

kp2: generate random variates
```{r}
library(tibble)
system.time(
  myVec <- rnorm(1000000,mean=10,sd=5)
)
time <- c(0.073,0.667,6.458,66.591)
sapSize <- c(1000000,10000000,100000000,1000000000)
sapSize.m <- sapSize/1000000
myDF <- tibble(sapSize.m,time)

library(tidyverse)

ggplot(data=myDF,aes(y=sapSize.m,x=time))+
  geom_line(linetype="dotted")+
  geom_point(size=4)+
  xlab("The running time")+
  ylab("The sample size of random variates in Millions")+
  scale_y_continuous(labels = comma)
```


kp3: R markdown options

create section / 
and header #,##,###,####,#####,######
To just show code out of code chunk, use 'a = 100'.
To execute code out of code chunk, use 'r sum(c(10,220))'

Chunk options:
include: T means: evaluate code, but do not show result and code.
Useful in load library, load data.

echo: T means show the code. F is used when you just need the output. 
Useful in making graph, giving summary statistics.

eval: T means to evaluate the code. F means just no do it.
Useful in just showing the code, may be used for comparison.

message: F prevent message showing.

cache: T store the result in the code chunk for future use.


```{r}
knitr::opts_chunk(echo=TRUE)
?type

```


kp4:show good table
```{r}
knitr::kable(dplyr::slice(mtcars,1:5))
```

kp5:type,class,mode decipher

type:
*typeof* determines the (R internal) type or storage mode of any object.

Current values are the vector types **"logical", "integer", "double", "complex", "character",** "raw" and "list", "NULL", "closure" (function), "special" and "builtin" (basic functions and operators), "environment", "S4" (some S4 objects)

5 common types: 
character,double,integer,logical,na

Coercion: logical -> integer -> double -> character

exp:typeof(100) is a double type. Strange.

you can also use is.atomic(someVec) to check about the type of vector. 

mode: 
Get or set the type or storage mode of an object.
types "integer" and "double" are returned as "numeric".

mode(x) <- "newmode" changes the mode of object x to newmode. This is only supported if there is an appropriate as.newmode function, for example "logical", "integer", "double", "complex", "raw", "character", "list", "expression", "name", "symbol" and "function".

'a' has class character,
100L has class numeric

```{r}
?type

vec <- seq(from=99,to = 96,by=-1)
typeof(vec[1])
typeof(100L)
mode('a')
mode(TRUE)
mode(0.4)
mode(100L)

myFavNum <- 3.1415
myNums <- c(1, 3, 55, 33, 86, -sqrt(2), -110) # also works myNums <- 1:7
firstNames <- c("Jeff", "Terence", "David") 
myVec <- c("Parr", 99)

d <- myFavNum+myNums[7]
#when you add a number to a atomic vector, each entry will get added.
dd <- myFavNum+myNums
(sum <- myFavNum+sum(myNums))
#the myVec will auto change type of data to most flexible, which will be character.
rs <- myVec[2]+myNums[2]
```
kp6:NA
```{r}
#create double NAs
NA_real_(10)
```

```{r}
#use order to arrange the vector, which will give the index of small value to large value.
myVec <- c(99.1, 98.2, 97.3, 96.4)
(order(myVec))
# Use the order sequence to print the ordered vector.
myVec[order(myVec)]
myVec[-c(-1)]
```
ex about vector
```{r}
myAt <- c(1, 4, 3, 2, NA, 3.22, -44, 2, NA, 0, 22, 34)
sum( myAt>0|myAt<0,na.rm=T)
```
kp7: list
```{r}
x1 <- list(c(1, 2), c(3, 4))
class(x1[1])
x2 <- list(list(1, 2), list(3, 4))
class(x2[1])
class(x1[[2]][1])#3
x1[[2]][1]#3 4
x1[[2]][[1]]#3
class(x1[2][1])#
```

kp8: factor 
based on integer vector, with two more attributes:
class() and levels(). The first make it behaves differently,
the latter define the kinds of data it can have.

```{r}
bodyType <- factor(c("healthy", "healthy", "healthy", "obese", "overweight", "overweight", "skinny"))
                  
str(bodyType)
```
Notice the levles healthy, obese,overweight,skinny is ordered by its appearance's order. And factor will assign an integer value w.r.t the level. 

But the integer comarison fails and we need to tell factor to use levels and sort by it. 
```{r}
bodyType <- factor(c("h", "h", "h", "ob", "ov", "ov", "s"), levels = c("s", "h", "ov", "ob"),
labels = c("Skinny", "Healthy", "Overweight", "Obese"), ordered = TRUE)

levels(bodyType)

bodyType[bodyType<"Overweight"]
```

Now the levels is mapped by skinny to 1, healthy to 2, so on.
And we can use name in labels to select.

```{r}
mycyl <- mtcars$cyl
mycyl
mycYl <- factor(mycyl,levels=c(4,6,8),labels=c("Small","Medium","Large"),ordered = T)
length(mycYl[mycYl<="Medium"])
```

```{r}
set.seed(4)
ageData <- round(runif(1000000, 0, 120))
groups <- cut(ageData,breaks=c(-Inf,2,5,9,12,19,65,Inf),levles=c("infant", "toddler", "child", "tween", "teen", "adult", "senior"),labels = c("infant", "toddler", "child", "tween", "teen", "adult", "senior"),ordered=T)
knitr::kable(groups)
```

```{r}
x <- matrix(1:10, ncol = 5, nrow = 2)
x
x[,1:3]

y <- array(1:12, c(2, 3, 2))
(y)
y[,,2]
```
```{r}
seqq <- 1:5
typeof(seqq)
?logical
logical(TRUE)
```


```{r}
df1 <- dplyr::starwars %>% select(height, homeworld) %>%
group_by(homeworld) %>%
summarise(meanHeight = mean(height, na.rm = T)) %>%
arrange(desc(meanHeight)) %>% ungroup() %>%
head(5)

df2 <- dplyr::starwars %>% select(height, homeworld) %>%
filter(homeworld %in% x[["homeworld"]])
```


kp: about group
```{r}
x <- as.character(1949:2017)
y <- runif(length(x),-20,20)
myDF <- data.frame(year=x,annualReturn = y)

ggplot(myDF)+geom_line(aes(x=year,y=annualReturn),group=1)
```
Compute the percentage of national population in each state. The result/output should look exactly like this, although it is truncated in length for ease of exposition):
```{r}
census <- read.csv("~/classes/msan593/data/2015census.csv")
library(magrittr)
library(dplyr)
census %>% group_by(State) %>% 
  summarise(totalPopu = sum(TotalPop)) %>% 
    mutate(popPerct = totalPopu/sum(totalPopu) * 100) %>% 
      arrange(desc(popPerct)) %>% 
        head(5) %>% 
  knitr::kable()

census1 <- census %>% select(State,TotalPop)
census1 %>% group_by(State) %>% summarise(tot=sum(TotalPop)) %>% arrange(desc(tot))
```



```{r}
data <- mtcars
str(data)

data1 <- data %>% group_by(cyl,gear) %>% summarize(n=n())
data1 %>% count(cyl,gear)
```




