---
title: "R_for_Data_Science"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r }
library(tidyverse)
library(magrittr)
df <- mpg
str(df)
```

## Including Plots

You can also embed plots, for example:

```{r }
ggplot(df)+
  geom_point(aes(x=drv,y=cyl))
```

```{r}
ggplot(df)+
  geom_point(aes(x= displ, y = hwy))+
  facet_wrap(~class,nrow=2)
```

```{r}
?geom_smooth
#exercies 3.6.1.6
p <- ggplot(data=df,aes(x=displ,y=hwy))
p+geom_point(size=4)+geom_smooth(se=F,color="blue")
p+geom_point(size=4)+geom_smooth(se=F,aes(group=drv))
p+geom_point(aes(color=drv))+geom_smooth(se=F,aes(color=drv))
p+geom_point(aes(color=drv))+geom_smooth(se=F)
p+geom_point(aes(color=drv))+geom_smooth(se=F,aes(linetype=drv))
p+geom_point(aes(color=drv),alpha=0.8)
```

EX 3.7.1
stat_summary operates on unique x

geom_bar makes the height of the bar proportional to the number of cases in each group.geom_bar uses stat_count by default: it counts the number of cases at each x position.

If you want the heights of the bars to represent values in the data, use geom_col instead.geom_col uses stat_identity: it leaves the data as is.
```{r}
?stat_summary
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
?geom_col

ggplot(data=df)+geom_col(aes(x=displ))
ggplot(data=df)+geom_bar(aes(x=displ))
ggplot(diamonds)+geom_boxplot(aes(x=cut,y=depth))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y =..prop..,fill=color))
```

```{r}
?geom_jitter
?geom_count
ggplot(data=df)+geom_jitter(aes(x=cty,y=hwy),width=0.35)
ggplot(data=df)+geom_count(aes(cty,hwy))
ggplot(data=df)+stat_sum(aes(cty,hwy))
```
```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```


```{r}
install.packages("map_data")
library(map_data)
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```

